---
title: "Final Project"
author: "Jose Luna"
date: "2022-11-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('readxl')
library(stats)
```
# Analyzing the Data.



```{r, echo=FALSE}
data =  read_excel('data.xlsx')
y = data$Anomaly
tm = 1:length(data$Anomaly)
plot(tm, y , type="l",xlab="Time", ylab="Anomalies")

```

 
We can see that the graph tends to increase over time and we can use time series to forecast the next values of the follwoing year. Its importaant to anylize with time series becasue we can decompose it with trends, if it either has cyclicallity and stationary



# Variation of the data.
We can see that our data foloows a normal distributio and it seen that the majority of the residuals are equally diustributed with some outliers at the begiing and the en of the data set

```{r, echo=FALSE}
# Graphical Techniques to understand the variability of that
loetrend = loess(y~tm, span = 0.5)
rougloess =loetrend$residuals
par(mfrow=c(1,2))
hist(rougloess, freq = F, main = "Histogram of Rough", xlab = "Rough part")
qqnorm(rougloess)
qqline(rougloess)
```
We can also do the test to see if the time series is statiomnat








# Stationary
in order to see if our data is station we need to take a look at the rough plot and also lag plot. 

```{r, echo=FALSE}

x = loetrend$residuals
plot(loetrend$residuals, type = "l", xlab = "Time", ylab = "Residuals", main = "Rough part")
lag.plot(x, lags = 1, do.lines = F)
```


```{r, echo=FALSE}

acf(x, lag.max = 10, main = "ACF plot of rough part")
pacf(x,main="PACF")
```
What does the ACF and PACF tells us about the data


